Appendix 3: Example of a spatially varying effect of climate on citizen science count data with sdmTMB

### 2021-12-29 doc/appendix-snow-sdmTMB.Rmd
more owl appendix edit
consider adding cv?

---
title: "Appendix 3: Example of a spatially varying effect of climate on citizen science count data with sdmTMB"
output:
  pdf_document:
    highlight: tango
    toc: true
    number_sections: true
urlcolor: blue
highlight: "monochrome"
bibliography: refs.bib
---

@@

# Introduction

@@

The result is a mesh with `r mesh$mesh$n` knots.

@@

@@

```{r check, echo=FALSE, include = FALSE, cache=FALSE, eval=TRUE}
if (rds_saved2) {
  m <- readRDS(here::here("data/snow_w_main_effect_150km_reml.rds"))
  m # loading this twice just prevents some excess messages from cluttering the output when printing below
}
```

```{r print-fit, warning=FALSE, echo=TRUE, message=FALSE, cache=FALSE}
print(m)
```

@@

We can also test for zero-inflation using the `simulate.sdmTMB` function and a couple functions from the DHARMa package.
If our data was zero-inflated we might needed to use a delta (aka. hurdle) model instead. 
This is not built in to sdmTMB, but can be achieve using the steps in the vignette [Fitting delta models with sdmTMB](https://https://pbs-assess.github.io/sdmTMB/articles/delta-models.html).

@@

>>>>

---
title: "Appendix 3: Example of a spatially varying effect of climate on citizen science count data with sdmTMB"
output:
  bookdown::pdf_document2:
    highlight: pygments
    toc: true
    number_sections: true
urlcolor: blue
bibliography: refs.bib
---

@@

# Background

@@

The result is a mesh with `r mesh$mesh$n` (`mesh$mesh$n`) knots.

@@

# Model validation

@@

```{r check, echo=FALSE, include = FALSE, eval=TRUE}
if (rds_saved2) {
  m <- readRDS(here::here("data/snow_w_main_effect_150km_reml.rds"))
  m # printing this first just prevents some excess messages from cluttering the output when printing below
}
```

```{r print-fit, warning=FALSE, echo=TRUE, message=FALSE}
print(m)
```

@@

We can also test for zero-inflation using the `simulate.sdmTMB` function and a couple functions from the DHARMa package.
If our data were zero-inflated or overdispersed, we might have needed to use a delta ("hurdle") model instead. 
This is not built directly into sdmTMB, but can be achieved by fitting the two components separately. See the steps in the vignette [Fitting delta models with sdmTMB](https://pbs-assess.github.io/sdmTMB/articles/delta-models.html).

@@

<!-- # ```{r} -->
<!-- # library(future) -->
<!-- # plan(multisession) -->
<!-- # m_cv <- sdmTMB_cv(count ~ 1 + nao + (1 | year_f), -->
<!-- #   spatial_varying = ~ 0 + nao, -->
<!-- #   time = "year",  -->
<!-- #   spatial = "on",  -->
<!-- #   spatiotemporal = "iid", -->
<!-- #   family = nbinom2(link = "log"), -->
<!-- #   # reml = TRUE, -->
<!-- #   use_initial_fit = TRUE, -->
<!-- #   data = snow, mesh = mesh, -->
<!-- #   k_folds = 10 -->
<!-- # ) -->
<!-- # m_cv -->
<!-- # ``` -->
